The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-10-06 rnw/Twophase_Estimators and their Application.Rnw
merge twophase 3

\begin{subequations}\label{eq:varexternal_2p_reg}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\hat{\var}(R(x))=\frac{1}{n_2}\frac{1}{n_{2}-1}\sum_{x \in {s_2}}\Big(\hat{R}(x)-\bar{\hat{R}}\Big)^2 \label{eq:varexternal_2p_reg_ex} \\
  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & = \frac{1}{n_1}\hat{\var}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}(R(x)) \nonumber \\ & =
 \frac{1}{n_1}\frac{1}{n_1-1}\sum_{x\in{s_1}}(\hat{Y}(x)-\bar{\hat{Y}}_{s_1})^2 + \frac{1}{n_2}\frac{1}{n_2-1}\sum_{x\in{s_2}}(\hat{R}(x)-\bar{\hat{R}})^2 \label{eq:varexternal_2p_reg_nex}
\end{align}
\end{subequations}

@@

One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}. In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case). Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e. deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$. While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large. Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks. One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support. Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible. In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$. This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association. With this respect, the infinite population approach can thus provide more flexibility than the finite approach. One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par

% 
% However, this is actually possible for some specific types of variables, such as when the auxiliary information is pixelated and the desired explanatory variable is the mean of pixel values in the support. However, for most types of explanatory variables we are going to try to get an approximation of $\bar{\pmb{Z}}$ that is only neglibly different. To do this we want \textit{all} of the wall-to-wall auxiliary information to be exploited. Figure \ref{fig:exh_nexh_and_boundweights_a} depicts such an implementation where the spatial arrangement of the supports (the dark green tiles) are tesselated to form a perfect partition over the inventory area. It has to be noted that this setup would allow for perfect calculation of the exact mean in the \textit{finite population approach} where the supports themselves would be considered the sampling units. While in the \textit{infinite population approach} this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large. In all other cases, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
% 
% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}. In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case). Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e. deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$. While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large. Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks. One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support. Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible. In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$. This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association. With this respect, the infinite population approach can thus provide more flexibility than the finite approach. One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par

% 
% 
% The perfect tesselation implementation may make it easier to calculate the exact auxiliary mean in the \textit{finite approach}, but it can also impose drawbacks. For example, the support size in practice should ideally be chosen in order to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) not to merely conform to the geometric constraints of a perfect tesselation.  Furthermore, the size of these supports are assumed to be the same for each explanatory variable.  This is not the case in the \textit{infinite population approach} where $\pmb{Z}(x)$ only refers to the sample point, not its spatial support. Thus, the supports are allowed to overlap if necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$ and each explanatory variable is allowed to have a different sized support as long as they are consistently used at all sample locations. In this respect, the \textit{infinite population approach} is more flexible.
% 
% It was mentioned that there are a few situations where the exact mean can be precisely calculated in the infinite population framework. The first common situation is when the explanatory variables are provided by polygon layers (e.g. map of development stages). In this case, one can calculate the exact mean as the area-weighted average of each categorical variable. The second case is when the exact mean can be calculated in one step, e.g. taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.
% 
% 
% % -----------------------
% %% my version:
% The \textbf{exhaustive} case however leaves some space for discussion concerning the calculation of $\pmb{Z}(x)$: This is because in the infinite population approach, the calculation $\pmb{Z}(x)$ again only refers to the sample point $x$ and has no spatial association. Likewise the mean of the local density (i.e. the total), the only way to determine the exact mean of $\pmb{Z}(x)$ would be to calculate the integral $\bar{\pmb{Z}}=\frac{1}{\lambda(F)}\int_{x \in F} \pmb{Z}(x)$, which in practice means to calculate $\pmb{Z}(x)$ at an infinite number of sample points $x$ ($n_1=\infty$). Since this is obviously impossible, one can only try to get a reasonable approximation of $\bar{\pmb{Z}}$.\par
% 
% One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}. In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case). Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e. deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$. While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large. Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks. One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support. Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible. In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$. This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association. With this respect, the infinite population approach can thus provide more flexibility than the finite approach. One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.\par
% Note that there are few cases where the exact mean can also be precisely calculated in the infinite population framework. The first case is when the explanatory variables are provided by polygon layers (e.g. map of development stages). In this case, one can calculate the exact mean as the area-weighted average of each categorical variable. The second case is when the exact mean can be calculated in one step, e.g. taking the mean of all height pixels of a raster canopy height model will perfectly equal the mean calculated by the use of an infinite number of supports \citep{mandallaz2013b}.

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(left).eps}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
   \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(right).eps}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_b}
	\end{subfigure}
\caption{Concept of exhaustive, see \textbf{(a)}, and non-exhaustive, see \textbf{(b)}, calculation of explanatory variables including boundary adjustment at the support level. Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle. A vector of explanatory variables $\pmb{Z}(x)$ is calculated within the supports (small squares)  at each sample location $x$ (points) that falls into the forest area (green underlying polygon). w: weights used for calculating the weighted mean of each explanatory variable. Dotted circles indicate terrestrial sample plots.}
\label{fig:exh_nexh_bw}
\end{figure}

@@

\begin{subequations}\label{eq:varext_2p_small_psmall}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2  \label{eq:varext_2p_reg_small} \\
  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(Y(x)-\bar{Y}_{2,G}\Big)^2 \nonumber \\
 &+ \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x \in s_{2,G}} \Big(\hat{R}(x)-\bar{\hat{R}}_{2,G}\Big)^2 \label{eq:varext_2p_reg_psmall}
\end{align}
\end{subequations}

If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.

@@

\begin{subequations}\label{eq:ext_varexternal_2p_extsynth}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}} \frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}})^2 \label{eq:ext_varexternal_2p_extsynth} \\
  \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & =
 \frac{1}{n_{1,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(Y(x)-\bar{Y}_{2,G})^2 \nonumber \\
 &+ (1-\frac{n_{2,G}}{n_{1,G}})\frac{1}{n_{2,G}}\frac{1}{n_{2,G}-1}\sum_{x\in{s_{2,G}}}(\hat{\mathbb{R}}(x)-\bar{\hat{\mathbb{R}}})^2 \label{eq:ext_varexternal_2p_extpsynth}
\end{align}
\end{subequations}

>>>>

\begin{subequations}\label{eq:varexternal_2p_reg}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{reg2p,ex}) & = \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \label{eq:varexternal_2p_reg_ex} \\
  \hat{\var}_{ext}(\hat{Y}_{reg2p,nex}) & = \frac{1}{n_1}\hat{\var}_{s_1}(\hat{Y}(x)) + \frac{1}{n_2}\hat{\var}_{s_2}(\hat{R}(x)) \nonumber  \label{eq:varexternal_2p_reg_nex}
\end{align}
\end{subequations}
where $\hat{\var}_{s_2}$ and $\hat{\var}_{s_1}$ indicate taking the sample variance over $s_2$ and $s_1$ respectively.

@@

One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Figure \ref{fig:exh_nexh_and_boundweights_a}, which illustrates the implemented solution to the exhaustive use of the auxiliary information described in \citet{mandallaz2013b}. In order for all of the wall-to-wall auxiliary information to be exploited, the spatial arrangement of the supports were tesselated to form a perfect partition over the inventory area as is the case with the dark green polygons (squares tiles in this case). Is has to be noted that this would perfectly enable for calculating the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e. deriving $\pmb{Z}(x)$ for the entire finite population of supports and consequenlty calculating the exact mean $\bar{\pmb{Z}}$. While this is not equal to the exact mean $\bar{\pmb{Z}}$ in the infinite population approach, $n_1$ is still expected to be reasonably large to assume a good approximation of $\bar{\pmb{Z}}$ as long as the size of the supports are not unreasonably large. Whereas this implementation provides a convenient flexibility to switch between extimators of both the finite and infinite approach using the \textit{same} data set, it can also impose practical drawbacks. One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ in combination with the spatial extent of the support. Since in practice the support size will rather be chosen in order to achieve a best possible explanatory power of the regression model, thus minimizing the residual variation, a perfect tesselation might not always be feasible. In the infinite population frame however, the supports are also allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ of $\pmb{Z}(x)$ to get a reasonable approximation of $\bar{\pmb{Z}}$. This is again because $\pmb{Z}(x)$ only refers to the sample point and has no spatial association. With this respect, the infinite population approach can thus provide more flexibility than the finite approach. One could even consider the use of different support sizes per explanatory variable as long as they are consistently used at all sample locations.

\begin{figure}[htb]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
    \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(left).eps}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_a}
		\end{subfigure}
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
   \resizebox{0.8\hsize}{!}{\includegraphics{fig/boundaryweight_graphic(right).eps}}%
		\caption{} \label{fig:exh_nexh_and_boundweights_b}
	\end{subfigure}
\caption{Concept of exhaustive, see \textbf{(a)}, and non-exhaustive, see \textbf{(b)}, calculation of explanatory variables including boundary adjustment at the support level. Auxiliary data are in both cases available over the entire inventory area marked by the large rectangle. A vector of explanatory variables $\pmb{Z}(x)$ is calculated within the supports (small squares)  at each sample location $x$ (points) that falls into the forest area (green underlying polygon).}
\label{fig:exh_nexh_bw}
\end{figure}

@@

\begin{subequations}\label{eq:varext_2p_small_psmall}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,small,2p}) & := \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{R}(x))  \label{eq:varext_2p_reg_small} \\
  \hat{\var}_{ext}(\hat{Y}_{G,psmall,2p}) & := \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{R}(x)) \label{eq:varext_2p_reg_psmall}
\end{align}
\end{subequations}
where $\hat{\var}_{s_{2,G}}$ indicates taking the sample variance over $s_{2,G}$. If boundary adjustment is applied, the simple mean of the explanatory variable over the small area $\hat{\bar{\pmb{Z}}}_G=\frac{1}{n_{1,G}}\sum_{x \in s_{1,G}}\pmb{Z}(x)$ is replaced by its weighted version $\hat{\bar{\pmb{Z}}}_G=\frac{\sum_{x\in{s_{1,G}}}w(x)\pmb{Z}(x)}{\sum_{x\in{s_{1,G}}}w(x)}$, and likewise for exhaustively used auxiliary information.

@@

\begin{subequations}\label{eq:ext_varexternal_2p_extsynth}
\begin{align}
  \hat{\var}_{ext}(\hat{Y}_{G,extsynth,2p}) & = \frac{1}{n_{2,G}}\hat{\var}_{s_{2,G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extsynth} \\
  \hat{\var}_{ext}(\hat{Y}_{G,extpsynth,2p}) & = \frac{1}{n_{1,G}}\hat{\var}_{s_{2,G}}(Y(x)) + \Big(1-\frac{n_{2,G}}{n_{1,G}}\Big)\frac{1}{n_{2,G}}\hat{\var}_{s_{2_G}}(\hat{\mathbb{R}}(x)) \label{eq:ext_varexternal_2p_extpsynth}
\end{align}
\end{subequations}
where $\hat{\mathbb{R}}(x)=Y(x)-\bar{\pmb{\mathbb{Z}}}^t_{G}\hat{\pmb{\theta}}_{s_2}$.
